Active contours-driven registration method for the structure-informed segmentation of diffusion MR images

### 2015-04-28 Publications/2015-NeuroImage/00-main.tex
Elsevier's proofreader edits

\title{Active-contours driven registration method for structure-informed segmentation of diffusion MR images}

@@

\address[bit]{Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, %
Universidad Polit\'ecnica de Madrid, Madrid, Spain}
\address[ciber]{Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Zaragoza, Spain}
\address[ucla]{Department of Mathematics, University of California,
Los Angeles (UCLA), Los Angeles, CA, US}
\address[scil]{Computer Science Department, Faculty of Science, Universit\'e de Sherbrooke, 2500 Boulevard Universit\'e, Sherbrooke, QC J1K 2R1, Canada}
\address[lts5]{Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland}
\address[chuv]{Dept. of Radiology, CIBM, University
Hospital Center (CHUV) and University of Lausanne (UNIL), Lausanne, Switzerland}

@@

% Note that keywords are not normally used for peerreview papers.
\begin{keyword}
diffusion MRI \sep cortical~parcellation \sep segmentation %
\sep nonlinear~registration \sep active~contours \sep susceptibility~distortion.
\end{keyword}

>>>>

\title{Active contours-driven registration method for the structure-informed segmentation of diffusion MR images}

@@

\address[bit]{Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, Universidad Polit\'ecnica de Madrid, Madrid, Spain}
\address[ciber]{Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Zaragoza, Spain}
\address[ucla]{Department of Mathematics, University of California,
Los Angeles (UCLA), Los Angeles, CA, US}
\address[scil]{Computer Science Department, Faculty of Science, Universit\'e de Sherbrooke, 2500 Boulevard Universit\'e, Sherbrooke, QC J1K 2R1, Canada}
\address[lts5]{Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland}
\address[chuv]{Dept. of Radiology, CIBM, University
Hospital Center (CHUV) and University of Lausanne (UNIL), Lausanne, Switzerland}

@@

% Note that keywords are not normally used for peerreview papers.
\begin{keyword}
active~contours \sep cortical~parcellation \sep diffusion MRI \sep nonlinear~registration \sep segmentation \sep susceptibility~distortion.
\end{keyword}

### 2015-04-28 Publications/2015-NeuroImage/03-introduction.tex
Elsevier's proofreader edits

% -*- root: 00-main.tex -*-
\section{Introduction}\label{sec:introduction}
The accurate delineation of \gls*{wm} in \gls*{dmri} and the fusion of prior
  anatomical information extracted from a \gls*{t1} image of the same subject
  are crucial in a range of applications based on tractography, such as
  the extraction of structural connectivity \citep{craddock_imaging_2013} or
  tract-based spatial statistics \citep{smith_tractbased_2006}.
However, segmenting \gls*{dmri} data precisely, and coregistration of anatomical
  images are difficult due to several limitations.
First, \gls{dmri} images suffer from partial voluming of several tissues in 
  voxels located at their interfaces due to the low resolution of images
  (generally around \isores[mm]{2.2}).
Second, \gls*{dmri} schemes probe the diffusion process within the brain in 
  many angles called \glspl*{dwi}, completed by one or more baseline (\emph{b0}) 
  volumes without directional gradients.
The extremely low \gls*{snr} and high dimensionality of \glspl*{dwi} disable their
  direct use in segmentation.
The low contrast between \gls*{gm} and \gls*{wm} in the \emph{b0} volumes is not suitable for 
  segmentation either.
Lastly, \gls*{dmri} images are acquired using \gls*{epi} to speed up acquisition
  at the cost of introducing geometrical distortion and signal degradation and
  destruction \citep{jezzard_correction_1995}.
The artifact has been proven to cause an important impact on the anatomy extracted
  from \gls*{dmri}, particularly in certain fiber bundles \citep{irfanoglu_effects_2012}.
These limitations preclude segmentation in native \gls*{dmri} space, including
  registration-based approaches using anatomical images as source of structural
  information.

Early attempts to delineate the \gls*{wm} proceeded by thresholding the 
  \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
  the \gls*{dmri} data.} map.
However, the mask and subsequent analyses substantially depend on the threshold set
  \citep{taoka_fractional_2009} owing to the large variability of \gls*{fa} and its
  decay with the increase of incoherence between fibers.
\cite{zhukov_level_2003} proposed active contours with edges represented
  by levelsets, evolving on directionally invariant scalar maps.
\cite{rousson_level_2004} successfully segmented the corpus callosum with
  region-based levelsets on the eigenvalues of tensors fitted on the
  \gls*{dmri} data.
A large body of work targeted the definition of appropriate features to cluster,
  such as the 5D representation proposed by \cite{jonasson_segmentation_2005}.
Other efforts include mixture models on sets of directionally invariant maps
  \citep{liu_brain_2007}, iterative \citep{hadjiprocopis_unbiased_2005} and
  hierarchical \citep{lu_segmentation_2008} clusterings,
  graph-cuts \citep{han_experimental_2009},
  and volume fraction modeling \citep{kumazawa_improvement_2013}.

To solve the segmentation problem through registration, \cite{saad_new_2009} 
  used Pearson correlation to perform linear alignment of the \gls*{t1} and 
  the \emph{b0}.
Similarly, an also linear-registration method is \emph{bbregister} \citep{greve_accurate_2009}.
It uses active contours with edges to look for intensity boundaries in the \emph{b0}
  image.
The active contours are initialized by surfaces extracted from the 
  \gls*{t1} using \emph{FreeSurfer} \citep{fischl_freesurfer_2012}:
  the \emph{pial} surface (exterior of cortical \gls*{gm}) and the \emph{white}
  surface (the \gls*{wm}/\gls*{gm} interface). 
Since the \emph{b0} image only presents a detectable frontier for the pial surface,
  \emph{bbregister} is limited to align the cortical layer in this
  application.
This tool has become the standard method for its proven robustness, despite the 
  distortions found in \gls*{dmri} are nonlinear.
To overcome this issue, \emph{bbregister} excludes from the boundary search those
  regions typically warped by the artifact.
As the distortion is not accounted for, it must be addressed separately.
Nonlinear registration has been successfully performed between \gls*{t2} and \emph{b0}
  images for their similarity, but uniquely as a way to correct for distortions
  \citep{kybic_unwarping_2000,studholme_accurate_2000,wu_comparison_2008,tao_variational_2009}.
Further registration of \gls*{t1} and \gls*{t2} images was still required to map the anatomical
  information (and the \gls*{wm} segmentation) into \gls*{dmri} space.

Therefore, given the described circumstances, a method performing simultaneously
  segmentation in native \gls*{dmri} space and registration of the corresponding \gls*{t1} image
  to the \gls*{dmri} data could provide the optimal solution.
Previously, joint segmentation and registration has been successfully applied to
  longitudinal object tracking \citep{paragios_level_2003} and atlas-based
  segmentation \citep{gorthi_active_2011}.
The most extended approach is integrating active contours in image registration 
  methods.
\cite{unal_coupled_2005} and later \cite{wang_joint_2006},
  improved an existing method \citep{yezzi_variational_2003} based on linear registration
  to the nonlinear case implementing a free-form deformation field.
\cite{droske_mumfordshah_2009} reviewed the existing techniques, and proposed two different
  approaches to apply the Mumford-Shah functional \citep{mumford_optimal_1989} in simultaneous
  registration and segmentation, through the propagation of the deformation field from
  the contours to the whole image definition.
Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
  registration method in 2D using levelsets and a nonlinear elasticity smoother on the
  displacement vector field, which preserves topology even with very large deformations.
Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
  levelset function for the registration of several active contours, in the application
  of atlas-based segmentation.

The hypothesis under this research is that the registration and segmentation
  problems in \gls*{dmri} can be solved simultaneously, increasing the geometrical
  accuracy of the process.
We propose a tool called \regseg{} to exploit the prior information of shapes
  extracted on \gls*{t1} images to register the anatomical reference
  to \gls*{dmri} space, implicitly segmenting these data.
The approach uses active contours without edges \citep{chan_active_2001} that evolve driving a
  free-form deformation field of B-Spline basis functions.
Optimization is performed using a descent strategy of explicit shape-gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
Hence, unlike the most of presented methods, \regseg{} does not implement levelsets.
Since the nonlinear distortion is aligned with one of the imaging axis (see 
  \autoref{sec:human_connectome}), \regseg{} includes an anisotropic regularizer of
  the displacements field proposed by \cite{nagel_investigation_1986}.
Finally, we evaluate \regseg{} using an extension of our instrumentation framework
  \citep{esteban_simulationbased_2014} that simulates known and realistic distortions
  on \gls*{dmri} data.
The evaluation includes a comparison to the \gls*{t2b} correction by integrating in the framework
  an in-house implementation of the method.

>>>>

% -*- root: 00-main.tex -*-
\section{Introduction}\label{sec:introduction}
The accurate delineation of \gls*{wm} in \gls*{dmri} and the fusion of prior
  anatomical information extracted from a \gls*{t1} image for the same subject
  are crucial in a range of applications based on tractography, such as
  the extraction of structural connectivity \citep{craddock_imaging_2013} or
  tract-based spatial statistics \citep{smith_tractbased_2006}.
However, the precise segmentation of \gls*{dmri} data and the coregistration of anatomical
  images are difficult due to several limitations.
First, \gls{dmri} images have a resolution that is much higher than that of the imaged
  microstructural features \citep{basser_microstructural_1996}.
Therefore, voxels located in structural discontinuities are affected by partial
  voluming of the signal sources.
Second, \gls*{dmri} schemes probe the diffusion process within the brain at
  many angles to obtain \glspl*{dwi}, which are completed by one or more baseline (\emph{b0})
  volumes without directional gradients.
The extremely low \gls*{snr} and the high dimensionality of \glspl*{dwi} prevent their
  direct use in segmentation.
The low contrast between \gls*{gm} and \gls*{wm} in the \emph{b0} volumes also makes them unsuitable for
  segmentation.
Finally, \gls*{dmri} images are acquired using \gls*{epi} to speed up the acquisition process but
  at the cost of introducing geometrical distortion, as well as signal degradation and
  destruction \citep{jezzard_correction_1995}.
The artifacts generated have major impacts on the anatomy extracted
  from \gls*{dmri}, particularly in certain fiber bundles \citep{irfanoglu_effects_2012}.
These limitations prevent segmentation in the native \gls*{dmri} space, including
  registration-based approaches using anatomical images as the sources of structural
  information.

Early attempts to delineate the \gls*{wm} involved thresholding the
  \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
  the \gls*{dmri} data.} map.
However, the mask and subsequent analyses are highly dependent on the threshold set
  \citep{taoka_fractional_2009}.
To overcome the unreliability of \gls*{fa} thresholding,
  \cite{zhukov_level_2003} proposed active contours with edges represented
  by level sets, which evolve on directionally invariant scalar maps.
\cite{rousson_level_2004} successfully segmented the corpus callosum with
  region-based level sets based on the eigenvalues of tensors fitted using the
  \gls*{dmri} data.
Many studies have addressed the definition of appropriate features for clustering,
  such as the 5D representation proposed by \cite{jonasson_segmentation_2005}.
Other efforts include mixed models based on sets of directionally invariant maps
  \citep{liu_brain_2007}, iterative \citep{hadjiprocopis_unbiased_2005} and
  hierarchical \citep{lu_segmentation_2008} clustering,
  graph-cuts \citep{han_experimental_2009},
  and volume fraction modeling \citep{kumazawa_improvement_2013}.

To address the segmentation problem by registration, \cite{saad_new_2009}
  used the Pearson’s correlation coefficient to obtain a linear alignment of the \gls*{t1} and
  the \emph{b0}.
Similarly, a linear registration method was employed by \emph{bbregister} \citep{greve_accurate_2009},
  which uses active contours with edges to search for intensity boundaries in the \emph{b0}
  image.
The active contours are initialized using surfaces extracted from the
  \gls*{t1} using \emph{FreeSurfer} \citep{fischl_freesurfer_2012}:
  the \emph{pial} surface (exterior of the cortical \gls*{gm}) and the \emph{white}
  surface (the \gls*{wm}/\gls*{gm} interface).
The \emph{b0} image only includes a detectable frontier for the pial surface, so
  \emph{bbregister} is limited to aligning the cortical layer in this
  application.
This tool has become the standard method because of its proven robustness, although the
  distortions found in \gls*{dmri} are nonlinear.
To overcome this issue, \emph{bbregister} excludes the
  regions that are typically warped by artifacts from the boundary search.
However, because the distortion is not considered, it must be addressed separately.
Nonlinear registration has been performed successfully between \gls*{t2} and \emph{b0}
  images based on their similarity as a unique method for correcting distortion
  \citep{kybic_unwarping_2000,studholme_accurate_2000,wu_comparison_2008,tao_variational_2009}.
However, further registration of the \gls*{t1} and \gls*{t2} images is still required to map the anatomical
  information (and for \gls*{wm} segmentation) into the \gls*{dmri} space.

Therefore, given these issues, a method that simultaneously performs
  segmentation in the native \gls*{dmri} space and registration of the corresponding \gls*{t1} image
  onto the \gls*{dmri} data could provide the optimal solution.
To the best of our knowledge, this strategy has never been proposed for the application described above.
Previously, joint segmentation and registration have been applied successfully to other problems
such as longitudinal object tracking \citep{paragios_level_2003} and atlas-based
  segmentation \citep{gorthi_active_2011}.
The most thorough approach integrates active contours during image registration
  methods.
\cite{unal_coupled_2005}, and later \cite{wang_joint_2006},
  improved an existing method \citep{yezzi_variational_2003} based on linear registration
  to the nonlinear case by implementing a free-form deformation field.
\cite{droske_mumfordshah_2009} reviewed the existing techniques and proposed two different
  approaches for applying the Mumford-Shah functional \citep{mumford_optimal_1989} during simultaneous
  registration and segmentation by propagating the deformation field from
  the contours onto the whole image definition.
Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
  registration method in 2D using level sets and a nonlinear elasticity smoother on the
  displacement vector field, which preserves the topology even with very high deformation.
Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
  level set function to register several active contours during the application
  of atlas-based segmentation.

The hypothesis tested in the present study is that registration and segmentation
  problems in \gls*{dmri} can be solved simultaneously, thereby increasing the geometrical
  accuracy of the process.
Thus, we propose a tool called \regseg{} that exploits the prior information of shapes
  extracted using \gls*{t1} images to register the anatomical reference
  to \gls*{dmri} space, thereby implicitly segmenting these data.
The proposed approach uses active contours without edges \citep{chan_active_2001}, which evolve to drive a
  free-form deformation field of B-spline basis functions.
Optimization is performed using a descent strategy with explicit shape gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
Therefore, unlike most previous methods, \regseg{} does not implement level sets.
The nonlinear distortion is aligned with one of the imaging axes (see
  \autoref{sec:human_connectome}), so \regseg{} includes an anisotropic regularizer for
  the displacements field proposed by \cite{nagel_investigation_1986}.
Finally, we evaluated \regseg{} using an extension of our instrumentation framework
  \citep{esteban_simulationbased_2014}, which simulates known and realistic distortions
  based on \gls*{dmri} data.
The evaluation included a comparison with \gls*{t2b} correction by integrating it within the framework
  of an in-house implementation of the method.

### 2015-04-28 Publications/2015-NeuroImage/06-discussion.tex
Elsevier's proofreader edits

\paragraph*{Accuracy tests}
The hypothesis underpinning our research was that image registration can be reliably performed
  by searching for homogeneous regions in a multispectral image that correspond to precise contours
  from an atlas, or extracted from other image (i.e. a different time step).
We proved that active contours without edges can be used to successfully drive a
  deformation supported by B-Spline basis functions using digital phantoms.
We randomly deformed four different phantom models mimicking three homogeneous regions,
  and use them to simulate \gls*{t1} and \gls*{t2} images at two resolution levels.
After registration with \regseg{} we measured the Hausdorff distance between the
  projected contours using the ground-truth warping and our estimation.
We concluded that errors are significantly lower than voxel resolution.
We also assessed the 95\% \gls*{ci}, yielding an aggregate interval of
  0.64 - 0.66 [mm] in the low resolution phantoms (2.0mm isotropic voxel) and
  0.34 - 0.38 [mm] in the high resolution phantoms (1.0mm isotropic).
Therefore, we also concluded that the error was bounded above by the half of the
  voxel spacing.

\paragraph*{Application on real data}
We designed \regseg{} as a method to segment \gls*{dmri} data exploiting the
  anatomical information extracted in the corresponding \gls*{t1} image of the subject.
Applications of whole-brain tractography \citep{smith_tractbased_2006,craddock_imaging_2013}
  usually solve the problem in a two-step approach.
First, images are corrected for nonlinear distortions using auxiliary acquisitions
  such as fieldmaps \citep{jezzard_correction_1995}, \gls*{t2} images \citep{kybic_unwarping_2000},
  etc.
Second, the segmentation is projected from a reference \gls*{t1} image using linear
  registration \citep{greve_accurate_2009}.
\Regseg{} solves this joint problem in one single step, and it does not require any additional
  acquisition over the minimal protocol including \gls*{t1} and \gls*{dmri} images only.
This situation is commonly found in historical datasets.

We evaluated \regseg{} in a real environment, using the experimental framework presented
  in \autoref{fig:evworkflows}.
We processed 16 subjects from the \gls*{hcp} database with both \regseg{}
  and an in-house replication of the \acrfull*{t2b} method.
\Regseg{} showed very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56-0.66 [mm],
  below the pixel size of 1.25mm.
The misregistration error remaining after \regseg{} was significantly lower ($p < 0.01$) than the
  error corresponding to the \gls*{t2b} correction, using Kruskal-Wallis H-tests
  (\autoref{tab:results_real}).
Visual inspection of all the results (\suppl{section S5}) and the violin plots included in
  \autoref{fig:results_real} confirmed that \regseg{} overperformed the \gls*{t2b} method
  in our settings.
Even though we carefully configured the \gls*{t2b} method by using the same algorithm and
  same settings of a widely used software, cross-comparison experiments are prone to
  the so-called \emph{instrumentation bias} \citep{tustison_instrumentation_2013}.
Therefore, these results do not prove that \regseg{} \emph{is better than} \gls*{t2b}.
The results reported let us propose \regseg{} as a solid option in the application field.
Additionally, the \gls*{t2b} may introduce an additional (and small) error in the necessary
  registration of \gls*{t2} to \gls*{t1} space.

>>>>

\paragraph*{Accuracy tests}
The hypothesis tested in our study was that reliable image registration can be performed
  by searching for homogeneous regions in a multispectral image, which correspond to precise contours
  from an atlas, or extracted from another image (i.e., a different time step).
We demonstrated that active contours without edges can be used successfully to drive a
  deformation supported by B-spline basis functions with digital phantoms.
We randomly deformed four different phantom models to mimic three homogeneous regions
  (\gls*{wm}, \gls*{gm}, and \acrlong*{csf}) and we used them to simulate \gls*{t1} and \gls*{t2}
  images at two resolution levels.
After registration with \regseg{}, we measured the Hausdorff distance between the
  projected contours obtained using the ground-truth warping and our estimates.
We concluded that the errors were significantly lower than the voxel resolution.
We also assessed the 95\% \gls*{ci}, which yielded an aggregate interval of
  0.64--0.66 [mm] for the low resolution phantoms (2.0 mm isotropic voxel) and
  0.34--0.38 [mm] for the high resolution phantoms (1.0 mm isotropic).
Therefore, we also concluded that the error was bounded above by half of the
  voxel spacing.

\paragraph*{Application to real data}
We designed \regseg{} as a method for segmenting \gls*{dmri} data by exploiting the
  anatomical information extracted from the corresponding \gls*{t1} image of the subject.
Applications in whole-brain tractography \citep{smith_tractbased_2006,craddock_imaging_2013}
  usually solve this problem with a two-step approach.
First, the images are corrected for nonlinear distortions using auxiliary acquisitions
  such as field maps \citep{jezzard_correction_1995}, \gls*{t2} images \citep{kybic_unwarping_2000}.
Second, the segmentation is projected from a reference \gls*{t1} image using linear
  registration \citep{greve_accurate_2009}.
\Regseg{} addresses this joint problem in a single step and it does not require any additional
  acquisition other than the minimal protocol using only \gls*{t1} and \gls*{dmri} images.
This situation is found commonly in historical datasets.

We evaluated \regseg{} in a real environment using the experimental framework presented
  in \autoref{fig:evworkflows}.
We processed 16 subjects from the \gls*{hcp} database using both \regseg{}
  and an in-house replication of the \acrfull*{t2b} method.
\Regseg{} obtained very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56--0.66 [mm], which was
  below the pixel size of 1.25 mm.
The misregistration error that remained after \regseg{} was significantly lower ($p < 0.01$) than the
  error corresponding to the \gls*{t2b} correction according to Kruskal-Wallis H-tests
  (\autoref{tab:results_real}).
Visual inspections of all the results (\suppl{section S5}) and the violin plots in
  \autoref{fig:results_real} confirmed that \regseg{} performed better than the \gls*{t2b} method
  in our settings.
We carefully configured the \gls*{t2b} method using the same algorithm and the 
  same settings employed widely in software, but cross-comparison experiments are prone to
  so-called \emph{instrumentation bias} \citep{tustison_instrumentation_2013}.
Therefore, these results do not prove that \regseg{} \emph{is better than} \gls*{t2b}.
However, our results suggest that \regseg{} is a reliable option in this application field.
In addition, the \gls*{t2b} may introduce an additional (and small) error during the necessary
  registration of \gls*{t2} in the \gls*{t1} space.

### 2015-04-28 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
major update
updates and standalone figures
Elsevier's proofreader edits

\subsection{Simplifying the regularization term}\label{app:reg_term}
The exponentials of the Thikonov regularization prior \eqref{eq:thikonov} have the general form
  $\vec{v}^T \mathbf{M} \vec{v}$.
If $\mathbf{M}$ is a $n \times n$ diagonal matrix such that $\mathbf{M} = \vec{m} \, \mathbf{I}_n$, 
  then:
\begin{equation*}
\vec{v}^T \mathbf{M} \vec{v} = \vec{m} \cdot (\vec{v}^T \mathbf{I}_n \vec{v}) = \vec{m} \cdot \vec{v}^{\circ2},
\end{equation*}
%
  where we have introduced the Hadamard power notation\footnote{The Hadamard power of a matrix or a vector
  is the power of its elements $\mathbf{M}^{\circ p} = ({m_{ij}}^{p})$}.

Since the anisotropy is generally aligned with the imaging axes, 
  $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:energy} can be simplified to diagonal matrices, such that
  $\mathbf{A}= \, \boldsymbol{\alpha}\,\vec{I}_n$ and
  $\mathbf{B}= \, \boldsymbol{\beta}\,\vec{I}_n$.
Substituting in the equation \eqref{eq:energy}, we get:
  \begin{align}
  E(M \mid U) &= \const + \underset{l}{\sum} \int_{\Omega_l}
  \mdist{f'}{l} \,d\vec{r} \notag\\
  &+ \int_{\Omega} \frac12 \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2} + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}.
  \label{eq:app_energy}
  \end{align}

>>>>

\subsection{Simplifying the regularization term}\label{app:reg_term}
The exponentials of the Thikonov regularization prior \eqref{eq:thikonov} have the general form
  $\vec{v}^T \mathbf{M} \vec{v}$.
If $\mathbf{M}$ is a $n \times n$ diagonal matrix such that $\mathbf{M} = \vec{m} \, \mathbf{I}_n$, 
  then:

\begin{equation*}
\vec{v}^T \mathbf{M} \vec{v} = \vec{m} \cdot (\vec{v}^T \mathbf{I}_n \vec{v}) = \vec{m} \cdot \vec{v}^{\circ2},
\end{equation*}
  where we have introduced the Hadamard power notation\footnote{The Hadamard power of a matrix or a vector
  is the power of its elements $\mathbf{M}^{\circ p} = ({m_{ij}}^{p})$}.

In general, the anisotropy is aligned with the imaging axes, so 
  $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:energy} can be simplified to diagonal matrices, such that
  $\mathbf{A}= \, \boldsymbol{\alpha}\,\vec{I}_n$ and
  $\mathbf{B}= \, \boldsymbol{\beta}\,\vec{I}_n$.
By substituting into equation \eqref{eq:energy}, we obtain:

  \begin{align}
  E(\vec{u}) &= \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} +   \int_{\Omega} \frac12 \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2} + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}.
  \label{eq:app_energy}
  \end{align}
